Primal-dual algorithm for computing Nash equilibria in sequential games

### 2014-12-12 nips2015.tex
more
wip

\documentclass{article} % For LaTeX2e
%\usepackage{nips14submit_e,times}
\usepackage{hyperref}
\usepackage{url}
\usepackage{amsmath,amsfonts,amsthm}

@@

\author{
DOHMATOB Elvis \inst[1]}
%% \institute{Parietal team, Inria Saclay Ile-de-France\\
%% Saclay, France\\

\title{A fast and simple primal-dual algorithm for computing Nash equilibria in two-person zero-sum games}

@@

\begin{abstract}
We present a simple $\mathcal{O}(1/\epsilon)$ primal-dual algorithm with a low cost per iteration, for computing Nash equilibria in two-person zero-sum games. Our algorithm is applicable to a broad class of two-person zero-sum games including simultaneous games and sequential games with imcomplete information and perfect recall. The applicability to the latter kind of games is thanks to the sequence-form representation von Stengel's \cite{von1996efficient}. Our algorithm derives from the primal-dual algorithm developed by Chambolle and Pock in \cite{chambolle2010}, which has gained considerable success in the signal processing community, but is not directly applicable to the saddle-point problem corresponding to Nash equilibria, since the strategy profiles of players are sufficiently complex
% (for example, simplexes for simultaneous games, and intersections of hyperplanes and nonnegative orthants, for sequential games)
 to prevent easy computation 
% (i.e using non-iterative procedures, etc.)
of euclidean
% (metric)%
 projections thereupon (the proximal operators).
Our technique is to dualize the constraints in the Nash equilibrium problem, thus effectively unconstraining the problem. The resulting algorithm is simple, only involving matrix-vector multiplications
% (without any matrix inversions, etc.)
and nonnegative clipping.
% (taking point-wise max of a vector and 0)
The $\mathcal{O}(1/\epsilon)$ convergence rate of our algorithm derives explicitly from the general primal-dual algorithm in \cite{chambolle2010}. As proof of concept, we apply our algorithm to solve Kuhn Poker.
\end{abstract}

\section{Introduction}
\label{sec:intro}
Though \textit{Nash equilibrium} has become a novel way for solving two-person zero-sum games, the problem of computing Nash equilibria for large-scale two-person zero-sum games is still largely unsolved. Even though this problem is, in theory, amendable to classical tools like LCP (\textit{Linear Constraint Programming}), at least for sequential games with imcomplete information the game (e.g Texas Hold'em Poker) is exceedingly larger than what such programs can handle.
%% We will be needing the following notation: %%  The reader should lookup any standard textbook
%% %% (for example \cite{boyd2004}) on convex optimization for a tutorial introduction to these notions.
%% Viz,
%% \begin{itemize}
%% \item $\mathbb{R}^n$: $n$-dimensional real vector space;
%% \item $\mathbb{R}^{m \times n}$: \quad space of all $m$-by-$n$ real matrices;
%% \item $0_{m,n}$: $m$-by-$n$ matrix of zeros;
%% \item $(x)_+$: \quad component-wise maximum of a vector $x$ and 0;
%% \item $\mathbb{R}^n_+$: \quad $\{x \in \mathbb{R}^n|x = (x)_+\}$, the $n$-dimensional nonnegative orthant 
%% \item $i_C$: \quad indicator function of a convex set $C$;
%% % \item $\Pi_C$: \quad euclidean projection operator onto a convex set $C$;
%% \item $\|K\|_2$: \quad spectral norm of a matrix $K$
%% % \item $F^*$: the convex conjugate of a convex function $F$.
%% %% \item \textit{l.s.c.p.c}: \quad acronym for adjective \textit{lower semi-continuous proper convex};
%% %% \item $f^*$: \quad Fenchel transform (a.k.a convex conjugate) of a \textit{l.s.c.p.c} function $f$;
%% \end{itemize}

\section{Statement of the problem}
We are interested in computing Nash equilibria for \textit{two-person zero-sum} games. The problem takes the well-known \textit{minimax} form
\begin{equation}
  \underset{y \in Q_2}{minimize}\text{ }\underset{x \in Q_1}{maximize}\text{ }{x^TAy}
  \label{eq:opt_pb}
\end{equation}

where \textit{feasibility sets} $Q_j$ have the form
\begin{equation}
  Q_j := \{z \in C_j|E_jz = e_j\}
\end{equation}
for some convex subset $C_j$ of a euclidean space $\mathbb{R}^{n_j}$, and some vector $e_j \in \mathbb{R}^{p_j}$.
$Q_j$ is the \textit{strategy portfolio} for player $j$. $A$ is the \textit{payoff matrix} from player 1's perspective of the game: if player 1 players strategy
$x \in Q_1$ and player 2 plays strategy $y \in Q_2$ then player 1 gets $x^TAy$ units of money; that the game is ``zero-sum'' means that player 2 gets $-x^TAy$.
We will assume that the convex sets $C_j$ are simple enough so that the eucliean projection operators $\Pi_{C_j}$ can be cheaply computed.
Typically, $C_j = \mathbb{R}^{p_j}_+ := \{x \in \mathbb{R}^n|x_j \ge 0\}$, the \textit{the nonnegative orthant}, and encodes a nonnegativity constraint. %% in this case, $\Pi_{C_j}(z) \equiv (z)_+$ as seen in subsection \ref{sec:notation}.
As usual, the ``minimax'' notation in problem \eqref{eq:opt_pb} means that a pair $(x^*, y^*) \in Q_1 \times Q_2$ is a solution iff
\begin{equation}
  x^TAy^* \le x^TAy \le {x^*}^TAy, \forall (x, y) \in Q_1 \times Q_2
\end{equation}
Such pairs $(x^*, y^*)$ correspond to the Nash equilibria of the game, and ${x^*}^TAy^*$ is the \textit{value}
\footnote{This value is the same for every equilibrium pair $(x^*, y^*)$.} of the game.

We can distinguish two categories of such games, in terms of (non-)simultaneity of play.
\paragraph{Simultaneous two-person zero-sum games}
\label{subsec:example_games}
Here, the Nash equilibrium problem takes the form \eqref{eq:opt_pb} with $p_j = 1$, $C_j = \mathbb{R}^{p_j}_+$, $E_j = 1_{1, n_j}$,
  and $e_j = 1$, so that $Q_j$ is simply the probabability $n_j$-simplex $\Delta_{n_j}$. Each point in $\Delta_{n_j}$ corresponds to a \textit{mixed-strategy} for player $j$,
and represents a randomization on their \textit{pure-strategies} (corresponding to the vertices of their propability simplex $\Delta_{n_j}$).

\paragraph{Two-person zero-sum sequential games with imcomplete information and perfect recall}
It is now known, thanks to the \textit{sequence-form representaion}, that the Nash equilibrium problem for such games takes the form \eqref{eq:opt_pb}.

We recall that in the sequence-form representation of such games,  $E_j$ is a matrix whose
entries are $-1$, $0$, or $+1$, and $e_j := (1, 0, 0, ..., 0)$. We also recall that $E_1$ and $e_1$ (resp. $E_2$ and $e_2$)
encode linear constraints player 1's (resp. player 2's)  ``admissible'' \textit{realization plans} $x$ (resp. $y$).

@@

\section{Some prerequisite notions from convex analysis}
\label{sec:notation}
We will need the following notations and definitions in the sequel. Given positive integers $m$ and $n$, $\mathbb{R}^{m \times n}$ denotes
the space of all $m$-by-$n$ real matrices. $0_{m,n}$ denotes the $m$-by-$n$ matrix of zeros and $1_{m,n}$ denotes the $m$-by-$n$ matrix of ones.
%% $\mathbb{R}^n_+$ := $\{x \in \mathbb{R}^n|x_j \ge 0 \text{ }  \forall j\}$ is the $n$-dimensional \textit{nonnegative orthant}.

@@

\section{The algorithm}
We now derive our primal-dual algorithm for solving the Nash equilibrium problem \eqref{eq:opt_pb}. The algorithm can is an application of the general primal-dual algorithm proposed in \cite{chambolle2010}. Though this scheme has recently gained considerable popularity in the signal processing community \cite{gramfort-etal:2013a, dohmatob2014benchmarking}, to the best of our knowledge, this is the first time it is being applied to compute Nash equilibria.

@@

  where $v \in \mathbb{R}^{p_1}$ and $u \in \mathbb{R}^{p_2}$ are auxiliary variables and 
  \begin{equation}
    \left .
    \begin{split}
      K :=
      \left[
        \begin{array}{c|c}
          A & -E_1^T \\ \hline
          E_2 & 0_{p_2, p_1}
        \end{array}
        \right] \in \mathbb{R}^{(n_1 + p_2) \times (n_2 + p_1)} \\
      %%\begin{bmatrix}A \text{ } E_1^T\\ E_2 \text{ } 0\end{bmatrix} \in \mathbb{R}^{(n_2 + p_1) \times (n_1 + p_2)}\\
      G: \mathbb{R}^{n_2} \times \mathbb{R}^{p_1} \rightarrow [0, +\infty], (y, v) \mapsto i_{C_2}(y) + e_1^Tv\\
      H: \mathbb{R}^{n_1} \times \mathbb{R}^{p_2} \rightarrow [0, +\infty], (x, u) \mapsto i_{C_1}(x) + e_2^Tu
    \end{split}
    \right\}
    \label{eq:things}
  \end{equation}

  Moreover, $G$ and $H$ are \textit{p.c.l.s.c} and their proximal operators are given by
  \begin{equation}
    \left .
    \begin{split}
      \text{prox}_{\tau G} : \mathbb{R}^{n_2} \times \mathbb{R}^{p_1} &\rightarrow \mathbb{R}^{n_2} \times \mathbb{R}^{p_1}\\
      (y, v) &\mapsto (\Pi_{C_2}(y), v - \tau e_1)\\
    \end{split}
    \right\}
  \end{equation}

  and
  \begin{equation}
    \left .
    \begin{split}
      \text{prox}_{\sigma F}: \mathbb{R}^{n_1} \times \mathbb{R}^{p_2} &\rightarrow \mathbb{R}^{n_1} \times \mathbb{R}^{p_2}\\
      (x, u) &\mapsto (\Pi_{C_1}(x), u - \sigma e_2)
    \end{split}
    \right\}
  \end{equation}
  \label{thm:pd}
\end{theorem}

%% Theorem \ref{thm:pd}, can be interpreted as follows: 2 players $x$ (player 1)  and $y$ (player 2) go heads-on in a zero-sum game with payoff matrix $A$. Their strategy, profiles are $Q_1$ and $Q_2$ respectively, meaning that if player $j$ players a strategy which is outside $Q_j$ then they are panalized by inflicting an infinite loss yupon them. To remedy this, they recruit assistants $u$ and $v$, whose sole goal is to preven the ...
\begin{proof} First observe that $\forall (x, y) \in \mathbb{R}^{n_1} \times \mathbb{R}^{n_2}$, we have
  \begin{eqnarray*}
      -i_{Q_1}(x) = -i_{C_1}(x) + \underset{v \in \mathbb{R}^{p_1}}{min}\text{}{v^T(e_1 - E_1x)},
      \end{eqnarray*}
  and
\begin{eqnarray*}
  i_{Q_2}(y) = i_{C_2}(y) + \underset{u \in \mathbb{R}^{p_2}}{max}\text{}{u^T(E_2y - e_2)}
\end{eqnarray*}
  Thus,% problem \eqref{eq:opt_pb} is equivalent to
  \begin{eqnarray*}
    \begin{split}
      x^TAy - i_{Q_1}(x) + i_{Q_2}(y) &= x^TAy -i_{C_1}(x) + \underset{v \in \mathbb{R}^{p_1}}{min}\text{}{v^T(e_1 - E_1x)} + i_{C_2}(y) + \underset{u \in \mathbb{R}^{p_2}}{max}\text{}{u^T(E_2y - e_2)}\\
      &= \underset{v \in \mathbb{R}^{p_1}}{min}\text{ }\underset{u \in \mathbb{R}^{p_2}}{min}\text{ }x^TAy - x^TE_1^Tv + u^TE_2y + i_{C_2}(y) + e_1^Tv -(i_{C_1}(x) + e_2^Tu)\\
      &= \underset{v\in \mathbb{R}^{p_1}}{min}\text{ }\underset{u \in \mathbb{R}^{p_2}}{max}
      {\begin{bmatrix}x\\u\end{bmatrix}^TK\begin{bmatrix}y\\v\end{bmatrix} + G(y, v) - H(x, u)},
    \end{split}
  \end{eqnarray*}
where $K$, $G$, and $H$ are as given in \eqref{eq:things}.

Hence,
\begin{eqnarray*}
  \begin{split}
    \underset{y \in Q_2}{min}\text{ }\underset{x \in Q_1}{max}\text{ }x^TAy &=
    \underset{y \in \mathbb{R}^{n_2}}{min}\text{ }\underset{x \in \mathbb{R}^{n_1}}{max}\text{ }x^TAy - i_{Q_1}(x) + i_{Q_2}(y)\\
    &= \underset{y \in \mathbb{R}^{n_2}}{min}\text{ }\underset{x \in \mathbb{R}^{n_1}}{max}\text{ }
\underset{v\in \mathbb{R}^{p_1}}{min}\text{ }\underset{u \in \mathbb{R}^{p_2}}{max}
      {\begin{bmatrix}x\\u\end{bmatrix}^TK\begin{bmatrix}y\\v\end{bmatrix} + G(y, v) - H(x, u)}\\
    &= \underset{y \in \mathbb{R}^{n_2}}{min}\text{ }\underset{v\in \mathbb{R}^{p_1}}{min}\text{ }\underset{x \in \mathbb{R}^{n_1}}{max}\text{ }
\underset{u \in \mathbb{R}^{p_2}}{max}
      {\begin{bmatrix}x\\u\end{bmatrix}^TK\begin{bmatrix}y\\v\end{bmatrix} + G(y, v) - H(x, u)}\\
      &=     \underset{y \in \mathbb{R}^{n_2}, v\in \mathbb{R}^{p_1}}{min}\text{ }\underset{x \in \mathbb{R}^{n_1}, u \in \mathbb{R}^{p_2}}{max}
           {\begin{bmatrix}x\\u\end{bmatrix}^TK\begin{bmatrix}y\\v\end{bmatrix} + G(y, v) - H(x, u)}
  \end{split}
\end{eqnarray*}
Thus problem \eqref{eq:opt_pb} is equivalent to problem \eqref{eq:unconstrained_pb} as claimed.

About the proximal operators, indeed both $G$ and $F$ are separable sums of the indicator function of a convex set, whose prox is simply the euclidean projection operator onto the set,  and a linear transformation $z \mapsto a^Tz$ whose prox, at rank $\tau$, is simply $z \mapsto z - \tau a$.
\end{proof}
Now applying Algorithm 39 of \cite{chambolle2010} with $F := H^*$ (the \textit{convex conjugate} of $H$)
 to problem \eqref{eq:unconstrained_pb}, we Algorithm \ref{Tab:algo} for solving the original problem \eqref{eq:opt_pb}. The $\mathcal{O}(1/\epsilon)$ convergence rate of Algorothm \ref{Tab:algo} hails directly from Theorem 1 of \cite{chambolle2010}. Note that this convergence rate cannot be improved in the framework of \cite{chambolle2010}, since neither $G$ nor $H$ is \textit{uniformly convex}.

>>>>

\documentclass{article} % For LaTeX2e
\usepackage{nips14submit_e,times}
\usepackage{hyperref}
\usepackage{url}
\usepackage{amsmath,amsfonts,amsthm}

@@

\author{
DOHMATOB Elvis}
%% \institute{Parietal team, Inria Saclay Ile-de-France\\
%% Saclay, France\\

\title{Primal-dual algorithm for computing Nash equilibria in
% two-person zero-sum
sequential games
% with imcomplete information and perfect recall
}

@@

\begin{abstract}
We present a simple $\mathcal{O}(1/\epsilon)$ primal-dual algorithm with a low cost per iteration, for computing Nash equilibria in two-person zero-sum sequential games like Texas Hold'em Poker. %% Our algorithm is applicable to a broad class of two-person zero-sum games including simultaneous games and sequential games with imcomplete information and perfect recall. The applicability to the latter kind of games is thanks to the sequence-form representation von Stengel's \cite{von1996efficient}
Our algorithm derives from the primal-dual algorithm developed
% by Chambolle and Pock
in \cite{chambolle2010}, which has received considerable interest in the signal processing community, but is not directly applicable to the saddle-point problem corresponding to Nash equilibria. This limitation is due to the fact that the strategy profiles (polytopes) of players are sufficiently complex
% (for example, simplexes for simultaneous games, and intersections of hyperplanes and nonnegative orthants, for sequential games)
 to prevent easy computation 
% (i.e using non-iterative procedures, etc.)
of euclidean
% (metric)%
 projections thereupon (the proximal operators).
Our technique is to dualize the linear constraints that form part of the player's strategy profile, thus obtainining an equivalent saddle-point problem which is more amendable to the primal-dual scheme. The resulting algorithm is simple, only involving matrix-vector multiplications
% (without any matrix inversions, etc.)
and nonnegative clipping.
% (taking point-wise max of a vector and 0)
The $\mathcal{O}(1/\epsilon)$ convergence rate of our algorithm derives explicitly from the general primal-dual algorithm in \cite{chambolle2010}. Unlike the EGT (Excessive Gap Technique) recently used in \cite{hoda2010smoothing}, our algorithm has a cheap and constant cost per iteration. As proof of concept, we apply our algorithm to solve Kuhn Poker.
\end{abstract}

\section{Introduction}
\label{sec:intro}
For sequential games with imcomplete information and perfect recall like Texas Hold'em Poker, von Stengel and co-workers have developed the \textit{sequence-form} representation (see \cite{von1996efficient} and references therewithin), whose size is linear in the size of the game tree, in contrast with the \textit{strategic-form} representation which is exponential in the size of the game tree.

%% It is now known, thanks to the \textit{sequence-form representaion}, that the Nash equilibrium problem for such games takes the form \eqref{eq:opt_pb}.

%%%
\paragraph{Sequence-form representaion}
The game consists of three players:
\begin{itemize}
\item player 1 (Alice),
\item player 2 (Bob), and
\item player 0, a special player called \textit{chance}, who plays a publicly known \textit{mixed strategy} $\beta_0$ (more on this later),
\end{itemize}
who take turns to play, according to the rules of the game. The sequence-form is similar to the \textit{extensive-form} except that in the latter, pure strategies are defined on edges, whilst they're defined on paths (sequences of moves) in the former.
The sequence-form representation for such a game is a quintuplet $(A, E_1, E_2, e_1, e_1)$, where
\begin{itemize}
\item For each \textit{information set} $h$ of player $k$, $\sigma_h$ is the sequence of moves to $h$. $C_h$ is the set of choices at $h$. $\mathcal{I}_k$ is the set of all information sets of player $k$, and the set of \textit{sequences} of player $k$, denoted $\mathcal{S}_k$, is defined by 
\begin{equation}
\mathcal{S}_k := \{\emptyset\} \cup \{\sigma_h c\text{ } |\text{ } h \in I_k, c \in C_h\}
\end{equation}
Each element $\sigma$ of $\mathcal{S}_k$ is a path joining the root node to a leaf in the game tree, with the moves of all other players (the other player, and the \textit{chance} player, aka \textit{nature}) deleted. Such a sequence represents a possible sequence of moves made by player $k$ in the game, from start to finish.
\item Given a node $t$
  \begin{itemize}
  \item $a(t)$ is the payoff received by player $1$ at $t$. Note that $a(t)$ is $0$ for all nonleaf nodes.
  \item $\sigma_k(t)$ is the unique sequence in $\mathcal{S}_k$ which can be extended to a path in the game tree, by including the moves of the other players;
  \item $\beta_0(t)$ is the product of the probabilities of the moves made by the chance player along $\sigma_0(t)$.
  \end{itemize}

\item Besides nonnegativity, the \textit{realization plans} $z$ of each play $k$ have linear contraints of the form
\begin{equation}
    z_{\emptyset} = 1 \text{ and }
    z_{\sigma_h} = \sum_{c \in C_h}{z_{\sigma_h c}}\text{, } \forall h \in \mathcal{I}_k
  \label{eq:lin}
\end{equation}

These linear constraints \eqref{eq:lin} can be written in the form
\begin{equation}
  E_k z = e_k
\end{equation}

where $E_k$ a matrix with $p_k := 1 + \#\mathcal{I}_k$ rows and $n_k := \#\mathcal{S}_k$ columns whose entries are either $-1$, $0$, or $1$, and $e_k$ is $p_k$-dimensional vector of the form $(1, 0, ..., 0)$. The \textit{strategy profile} of player $k$ is then the $n_k$-dimensional polytope
\begin{equation}
  Q_k := \{z \in \mathbb{R}^{n_k}| z \ge 0\text{ and }E_kz = e_k\}
  \label{eq:poly}
\end{equation}

Player $k$'s \textit{pure} strategies correspond to their sequences $\mathcal{S}_k$, and a realization plan is simply a randomization on these pure strategies.

\item $A$, an $n_1$-by-$n_2$ matrix, is the \textit{payoff matrix} from player 1's perspective of the game. Formally $A = (A_{\sigma,\tau})$, where
  \begin{equation}
    A_{\sigma,\tau} := \sum_{\text{leafs }t\text{ : } \sigma_1(t) = \sigma\text{ and } \sigma_2(t) = \tau}{\beta_0(t)a(t)}, \forall (\sigma, \tau) \in \mathcal{S}_1 \times \mathcal{S}_2
  \end{equation}

If player 1 plays strategy $x \in Q_1$ and player 2 plays strategy $y \in Q_2$ then player 1 gets $x^TAy$ units of money; that the game is ``zero-sum'' means that player 2 gets $-x^TAy$.

\end{itemize}

\paragraph{Nash equilibrium problem in sequence-form representaion}
In sequence-form representation, the Nash equilibrium problem is
\begin{equation}
  \underset{y \in Q_2}{minimize}\text{ }\underset{x \in Q_1}{maximize}\text{ }{x^TAy}
  \label{eq:opt_pb}
\end{equation}
As usual, the ``minimax'' notation in problem \eqref{eq:opt_pb} means that a pair $(x^*, y^*) \in Q_1 \times Q_2$ is a solution iff
\begin{equation}
  x^TAy^* \le x^TAy \le {x^*}^TAy, \forall (x, y) \in Q_1 \times Q_2
\end{equation}
Such pairs $(x^*, y^*)$ correspond to the Nash equilibria of the game, and ${x^*}^TAy^*$ is the \textit{value}
\footnote{This value is the same for every equilibrium pair $(x^*, y^*)$.} of the game.

Even though problem \eqref{eq:opt_pb} is, in theory, amendable to classical tools like LCP (\textit{Linear Constraint Programming}), at least for sequential games with imcomplete information the game (e.g Texas Hold'em Poker) is exceedingly larger than what such programs can handle.

%% Though \textit{Nash equilibrium} has become a novel way for solving two-person zero-sum games, the problem of computing Nash equilibria for large-scale two-person zero-sum games is still largely unsolved. Even though this problem is, in theory, amendable to classical tools like LCP (\textit{Linear Constraint Programming}), at least for sequential games with imcomplete information the game (e.g Texas Hold'em Poker) is exceedingly larger than what such programs can handle.
%% We will be needing the following notation: %%  The reader should lookup any standard textbook
%% %% (for example \cite{boyd2004}) on convex optimization for a tutorial introduction to these notions.
%% Viz,
%% \begin{itemize}
%% \item $\mathbb{R}^n$: $n$-dimensional real vector space;
%% \item $\mathbb{R}^{m \times n}$: \quad space of all $m$-by-$n$ real matrices;
%% \item $0_{m,n}$: $m$-by-$n$ matrix of zeros;
%% \item $(x)_+$: \quad component-wise maximum of a vector $x$ and 0;
%% \item $\mathbb{R}^n_+$: \quad $\{x \in \mathbb{R}^n|x = (x)_+\}$, the $n$-dimensional nonnegative orthant 
%% \item $i_C$: \quad indicator function of a convex set $C$;
%% % \item $\Pi_C$: \quad euclidean projection operator onto a convex set $C$;
%% \item $\|K\|_2$: \quad spectral norm of a matrix $K$
%% % \item $F^*$: the convex conjugate of a convex function $F$.
%% %% \item \textit{l.s.c.p.c}: \quad acronym for adjective \textit{lower semi-continuous proper convex};
%% %% \item $f^*$: \quad Fenchel transform (a.k.a convex conjugate) of a \textit{l.s.c.p.c} function $f$;
%% \end{itemize}

%% \section{Statement of the problem}
%% The Nash equilibrium problem for a two-person zero-sum sequential game with imcomplete information and perfect recall is a the saddle-point problem
%% \begin{equation}
%%   \underset{y \in Q_2}{minimize}\text{ }\underset{x \in Q_1}{maximize}\text{ }{x^TAy}
%%   \label{eq:opt_pb}
%% \end{equation}

%% where player $k$'s \textit{strategy profile} $Q_k$ is the polytope
%% \begin{equation}
%%   Q_k := \{z \in \mathbb{R}^{n_k}| E_kz = e_k\text{ and } z_j \ge 0 \text{ }\forall j\}
%% \end{equation}
%% for %% some convex subset $C_k$ of a euclidean space $\mathbb{R}^{n_k}$, and 
%% some $p_k$-by-$n_k$ matrix $E_k$ and $p_k$-dimensional vector $e_k$. $A$ is the \textit{payoff matrix} from player 1's perspective of the game: if player 1 players strategy
%% $x \in Q_1$ and player 2 plays strategy $y \in Q_2$ then player 1 gets $x^TAy$ units of money; that the game is ``zero-sum'' means that player 2 gets $-x^TAy$.
%% %% We will assume that the convex sets $C_k$ are simple enough so that the eucliean projection operators $\Pi_{C_k}$ can be cheaply computed.
%% Typically, $C_k = \mathbb{R}^{p_k}_+ := \{x \in \mathbb{R}^n|x_k \ge 0\}$, the \textit{the nonnegative orthant}, and encodes a nonnegativity constraint. %% in this case, $\Pi_{C_k}(z) \equiv (z)_+$ as seen in subsection \ref{sec:notation}.
%% As usual, the ``minimax'' notation in problem \eqref{eq:opt_pb} means that a pair $(x^*, y^*) \in Q_1 \times Q_2$ is a solution iff
%% \begin{equation}
%%   x^TAy^* \le x^TAy \le {x^*}^TAy, \forall (x, y) \in Q_1 \times Q_2
%% \end{equation}
%% Such pairs $(x^*, y^*)$ correspond to the Nash equilibria of the game, and ${x^*}^TAy^*$ is the \textit{value}
%% \footnote{This value is the same for every equilibrium pair $(x^*, y^*)$.} of the game.

%% We can distinguish two categories of such games, in terms of (non-)simultaneity of play.
%% \paragraph{Simultaneous two-person zero-sum games}
%% \label{subsec:example_games}
%% Here, the Nash equilibrium problem takes the form \eqref{eq:opt_pb} with $p_k = 1$, $C_k = \mathbb{R}^{p_k}_+$, $E_k = 1_{1, n_k}$,
%%   and $e_k = 1$, so that $Q_k$ is simply the probabability $n_k$-simplex $\Delta_{n_k}$. Each point in $\Delta_{n_k}$ corresponds to a \textit{mixed-strategy} for player $j$,
%% and represents a randomization on their \textit{pure-strategies} (corresponding to the vertices of their propability simplex $\Delta_{n_k}$).

%% \paragraph{Two-person zero-sum sequential games with imcomplete information and perfect recall}
%% It is now known, thanks to the \textit{sequence-form representaion}, that the Nash equilibrium problem for such games takes the form \eqref{eq:opt_pb}.

%% We recall that in the sequence-form representation of such games,  $E_k$ is a matrix whose
%% entries are $-1$, $0$, or $+1$, and $e_k := (1, 0, 0, ..., 0)$. We also recall that $E_1$ and $e_1$ (resp. $E_2$ and $e_2$)
%% encode linear constraints player 1's (resp. player 2's)  ``admissible'' \textit{realization plans} $x$ (resp. $y$).

@@

\section{Useful convex analysis}
\label{sec:notation}
We will need the following notations and definitions in the sequel. Given positive integers $m$ and $n$, $\mathbb{R}^{m \times n}$ denotes
the space of all $m$-by-$n$ real matrices. $0_{m,n}$ denotes the $m$-by-$n$ matrix of zeros and $1_{m,n}$ denotes the $m$-by-$n$ matrix of ones.
%% $\mathbb{R}^n_+$ := $\{x \in \mathbb{R}^n|x_k \ge 0 \text{ }  \forall j\}$ is the $n$-dimensional \textit{nonnegative orthant}.

@@

\section{The algorithm}
We now derive our primal-dual algorithm for solving the Nash equilibrium problem \eqref{eq:opt_pb}. The algorithm is derived from the general primal-dual Algorithm 39 proposed in \cite{chambolle2010}. Though this scheme has recently gained considerable popularity in the signal processing community \cite{gramfort-etal:2013a}, to the best of our knowledge, this is the first time it is being applied to compute Nash equilibria.

It should be noted that the difficulty of applying the general primal-dual Algorithm 39 proposed in \cite{chambolle2010} directly to problem \eqref{eq:opt_pb} lies in the difficulty of effectively computing the metric projections $\Pi_{Q_k}$. Theorem \ref{thm:pd} provides a trick to effectively evade this difficulty.

@@

  where $v \in \mathbb{R}^{p_1}$ and $u \in \mathbb{R}^{p_2}$ are auxiliary variables and 
  \begin{equation}
    \left .
    \begin{split}
      K :=
      \left[
        \begin{array}{c|c}
          A & -E_1^T \\ \hline
          E_2 & 0_{p_2, p_1}
        \end{array}
        \right] \in \mathbb{R}^{(n_1 + p_2) \times (n_2 + p_1)} \\
      %%\begin{bmatrix}A \text{ } E_1^T\\ E_2 \text{ } 0\end{bmatrix} \in \mathbb{R}^{(n_2 + p_1) \times (n_1 + p_2)}\\
      G: \mathbb{R}^{n_2} \times \mathbb{R}^{p_1} \rightarrow [0, +\infty], (y, v) \mapsto i_{\mathbb{R}^{n_2}_+}(y) + e_1^Tv\\
      H: \mathbb{R}^{n_1} \times \mathbb{R}^{p_2} \rightarrow [0, +\infty], (x, u) \mapsto i_{\mathbb{R}^{n_1}_+}(x) + e_2^Tu
    \end{split}
    \right\}
    \label{eq:things}
  \end{equation}

  Moreover, $G$ and $H$ are \textit{p.c.l.s.c} and their proximal operators are given by
  \begin{equation}
    \left .
    \begin{split}
      \text{prox}_{\tau G} : \mathbb{R}^{n_2} \times \mathbb{R}^{p_1} &\rightarrow \mathbb{R}^{n_2} \times \mathbb{R}^{p_1}\\
      (y, v) &\mapsto ((y)_+, v - \tau e_1)\\
    \end{split}
    \right\}
  \end{equation}

  and
  \begin{equation}
    \left .
    \begin{split}
      \text{prox}_{\sigma F}: \mathbb{R}^{n_1} \times \mathbb{R}^{p_2} &\rightarrow \mathbb{R}^{n_1} \times \mathbb{R}^{p_2}\\
      (x, u) &\mapsto ((x)_+, u - \sigma e_2)
    \end{split}
    \right\}
  \end{equation}
  \label{thm:pd}
\end{theorem}

%% Theorem \ref{thm:pd}, can be interpreted as follows: 2 players $x$ (player 1)  and $y$ (player 2) go heads-on in a zero-sum game with payoff matrix $A$. Their strategy, profiles are $Q_1$ and $Q_2$ respectively, meaning that if player $j$ players a strategy which is outside $Q_k$ then they are panalized by inflicting an infinite loss yupon them. To remedy this, they recruit assistants $u$ and $v$, whose sole goal is to preven the ...
\begin{proof} First observe that $\forall (x, y) \in \mathbb{R}^{n_1} \times \mathbb{R}^{n_2}$, we have
  \begin{eqnarray*}
   -i_{Q_1}(x) = -i_{\mathbb{R}^{n_1}_+}(x) + \underset{v \in \mathbb{R}^{p_1}}{min}\text{}{v^T(e_1 - E_1x)},
      \end{eqnarray*}
  and
\begin{eqnarray*}
  i_{Q_2}(y) = i_{\mathbb{R}^{n_2}_+}(y) + \underset{u \in \mathbb{R}^{p_2}}{max}\text{}{u^T(E_2y - e_2)}
\end{eqnarray*}
  Thus,% problem \eqref{eq:opt_pb} is equivalent to
  \begin{eqnarray*}
    \begin{split}
  \underset{y \in Q_2}{min}\text{ }\underset{x \in Q_1}{max}\text{ }{x^TAy} &=
  \underset{y \in \mathbb{R}^{n_2}}{min}\text{ }\left(\underset{x \in \mathbb{R}^{n_1}}{max}\text{ }
x^TAy - i_{Q_1}(x)\right) + i_{Q_2}(y) \\
&= \underset{y \in \mathbb{R}^{n_2}}{min}\text{ }\underset{x \in \mathbb{R}^{n_1}}{max}\text{ }x^TAy -i_{\mathbb{R}^{n_1}_+}(x) + \underset{v \in \mathbb{R}^{p_1}}{min}\text{}{v^T(e_1 - E_1x)} + i_{\mathbb{R}^{n_2}_+}(y) + \underset{u \in \mathbb{R}^{p_2}}{max}\text{}{u^T(E_2y - e_2)}\\
      &= \underset{y \in \mathbb{R}^{n_2}}{min}\text{ }\underset{x \in \mathbb{R}^{n_1}}{max}\text{ }\underset{v \in \mathbb{R}^{p_1}}{min}\text{ }\underset{u \in \mathbb{R}^{p_2}}{min}\text{ }x^TAy - x^TE_1^Tv + u^TE_2y + i_{\mathbb{R}^{n_2}_+}(y) + e_1^Tv -(i_{\mathbb{R}^{n_1}_+}(x) + e_2^Tu)\\
      &= \underset{y \in \mathbb{R}^{n_2}}{min}\text{ }\underset{x \in \mathbb{R}^{n_1}}{max}\text{ } \underset{v\in \mathbb{R}^{p_1}}{min}\text{ }\underset{u \in \mathbb{R}^{p_2}}{max}
      {\begin{bmatrix}x\\u\end{bmatrix}^TK\begin{bmatrix}y\\v\end{bmatrix} + G(y, v) - H(x, u)}\\
      &= \underset{y \in \mathbb{R}^{n_2}}{min}\text{ }\underset{v\in \mathbb{R}^{p_1}}{min}\text{ }\underset{x \in \mathbb{R}^{n_1}}{max}\text{ }
\underset{u \in \mathbb{R}^{p_2}}{max}
      {\begin{bmatrix}x\\u\end{bmatrix}^TK\begin{bmatrix}y\\v\end{bmatrix} + G(y, v) - H(x, u)}\\
      &=     \underset{y \in \mathbb{R}^{n_2}, v\in \mathbb{R}^{p_1}}{min}\text{ }\underset{x \in \mathbb{R}^{n_1}, u \in \mathbb{R}^{p_2}}{max}
           {\begin{bmatrix}x\\u\end{bmatrix}^TK\begin{bmatrix}y\\v\end{bmatrix} + G(y, v) - H(x, u)}
    \end{split}
  \end{eqnarray*}
where $K$, $G$, and $H$ are as given in \eqref{eq:things}.

Thus problem \eqref{eq:opt_pb} is equivalent to problem \eqref{eq:unconstrained_pb} as claimed.

About the proximal operators, indeed both $G$ and $F$ are separable sums of the indicator function of a convex set, whose prox is simply the euclidean projection operator onto the set,  and a linear transformation $z \mapsto a^Tz$ whose prox, at rank $\tau$, is simply $z \mapsto z - \tau a$.
\end{proof}
Now applying Algorithm 39 of \cite{chambolle2010} with $F := H^*$ (the \textit{convex conjugate} of $H$)
 to problem \eqref{eq:unconstrained_pb}, we obtain Algorithm \ref{Tab:algo_simplified} for solving the original problem \eqref{eq:opt_pb}. The $\mathcal{O}(1/\epsilon)$ convergence rate of Algorothm \ref{Tab:algo_simplified} hails directly from Theorem 1 of \cite{chambolle2010}. Note that this convergence rate cannot be improved in the framework of \cite{chambolle2010}, since neither $G$ nor $H$ is \textit{uniformly convex}.
